### 2018-07-12 source/Ch0-Foreword/foreword.tex
added various bits that came about after reshuffling the organisation

\pg
Radio astronomy has a short but storied history: from Karl Jansky's serendipitous observation of the centre of Sagittarius A*, which outshines our Sun in the radio regime, in 1933, to Grote Reber's hand-built back-yard radio antenna in 1937, which successfully detected radio emission from the Milky Way itself\footnote{\url{https://www.nrao.edu/whatisra/hist_reber.shtml}}, to such monumental projects as the Square Kilometer Array and its multiple pathfinders, it has led to countless discoveries and the opening of a truly new window on the Universe. The work presented in this thesis is a contribution to this discipline - the culmination of three years of study, which is a rather short time to get a firm grasp of radio interferometry both in theory and in practice. The need for robust, automated methods - which are improving daily, thanks to the tireless labour of the scientists in the field - is becoming ever stronger as the SKA approaches, looming large on the horizon; but even today, in the precursor era of LOFAR, MeerKAT and other pathfinders, it is keenly felt. When I started my doctorate, the sheer scale of the task at hand felt overwhelming - to actually be able to contribute to its resolution seemed daunting indeed!

>>>>

\pg
Radio astronomy has a short but storied history: from Karl Jansky's serendipitous observation of the centre of the Milky Way, which outshines our Sun in the radio regime, in 1933, to Grote Reber's hand-built back-yard radio antenna in 1937, which successfully detected radio emission from the Milky Way itself\footnote{\url{https://www.nrao.edu/whatisra/hist_reber.shtml}}, to such monumental projects as the Square Kilometer Array and its multiple pathfinders, it has led to countless discoveries and the opening of a truly new window on the Universe. The work presented in this thesis is a contribution to this discipline - the culmination of three years of study, which is a rather short time to get a firm grasp of radio interferometry both in theory and in practice. The need for robust, automated methods - which are improving daily, thanks to the tireless labour of the scientists in the field - is becoming ever stronger as the SKA approaches, looming large on the horizon; but even today, in the precursor era of LOFAR, MeerKAT and other pathfinders, it is keenly felt. When I started my doctorate, the sheer scale of the task at hand felt overwhelming - to actually be able to contribute to its resolution seemed daunting indeed!

### 2018-07-12 source/Ch1-ThisThesis/science.tex
added various bits that came about after reshuffling the organisation

\pg
Synchrotron radiation (or "magnetobremsstrahlung") occurs when the trajectory of a relativistic charged particle is deflected by a magnetic field. As the German name suggests, it is the magnetic equivalent of free-free radiation. It is the tracer of very energetic processes, such as the interaction between AGN jets and the intergalactic medium. In the low-energy regime, it is also known as cyclotron radiation, after the device in which it was first measured in a laboratory. %, and in non-relativistic regimes, it is known as gyro radiation.

@@

\pg
In the optically-thick regime, we know that the most efficient radiation mechanism is simple black-body radiation, and so the power spectrum at a given frequency can be defined in terms of an ``electron temperature". In the Rayleigh-Jeans approxmation (valid as we are explicitly in the low-frequency regime), this gives $S_\nu \propto \nu{5/2}$, or $\alpha=5/2$. 
In the optically-thin regime, meanwhile, we recover a mixture of the underlying power-law distribution ($S_\nu \propto e^{-\nu}$) and black-body radiation (which has a frequency-dependent $\alpha$). Empirically, we generally observe $\alpha \sim 0.7$.

>>>>

\pg
Synchrotron radiation (or ``magnetobremsstrahlung") occurs when the trajectory of a relativistic charged particle is deflected by a magnetic field. As the German name suggests, it is the magnetic equivalent of free-free radiation. It is the tracer of very energetic processes, such as the interaction between AGN jets and the intergalactic medium. In the low-energy regime, it is also known as cyclotron radiation, after the device in which it was first measured in a laboratory. %, and in non-relativistic regimes, it is known as gyro radiation.

@@

\pg
In the optically-thick regime, we know that the most efficient radiation mechanism is simple black-body radiation, and so the power spectrum at a given frequency can be defined in terms of an ``electron temperature". In the Rayleigh-Jeans approxmation (valid as we are explicitly in the low-frequency regime), this gives $S_\nu \propto \nu^{5/2}$, or $\alpha=5/2$. 
In the optically-thin regime, meanwhile, we recover a mixture of the underlying power-law distribution ($S_\nu \propto e^{-\nu}$) and black-body radiation (which has a frequency-dependent $\alpha$). Empirically, we generally observe $\alpha \sim 0.7$.

### 2018-07-12 source/Ch2-techniques/RIME.tex
added various bits that came about after reshuffling the organisation

\chapter{Mathematical Framework: the RIME Formalism}\label{section.RIME}
\minitoc

@@

\pg
Consider a single quasi-monochromatic point source in the sky. Its signal at a point in time and space can be described by a complex vector $\evector$. We can then represent physical processes which affect this signal's propagation using \emph{Jones matrices}. In other words:
\begin{equation}
\volt = \Jones\evector
\end{equation}
where $\volt$ is the voltage measured by our antenna, and $\Jones$ the Jones matrix describing the \emph{net} propagation effects - atmospheric, astrophysical, and instrumental - which affect our source's signal. $\Jones$ can be written as a series of matrix products, each individual matrix describing a physical phenomenon. They are then called a \emph{Jones chain}, and the resulting $\Jones$ is often called the \emph{total Jones matrix}.

\pg
A visibility is the correlation of voltage from two antennas. In other words,
\begin{align}
\Vis_{pq} &= 2 \volt_p (\volt_q)^\Herm\\
\Vis_{pq} &= 2 \Jones_p \evector (\Jones_q \evector)^\Herm\\
\Vis_{pq} &= 2 \Jones_p \evector \evector^\Herm \Jones_q^\Herm
\end{align}
where $\Jones_p$ corresponds to the signal propagation chain of antenna $p$, and $\Jones_q$ for antenna $q$. Note that a factor of 2 has been introduced here, as a matter of convention. This is due to the definition of the Stokes parameter in relation to our $\evector$ outer product.

\pg
If we choose to use $xyz$ as our coordinates, with $z$ the direction of propagation of our signal, then \citetads{1996A&AS..117..137H} show that the following relation holds:
\begin{equation}
2 \evector \evector^\Herm = 2  \begin{bmatrix} <e_x e_x^\star> & <e_x e_y^\star> \\ <e_y e_x^\star> & <e_y e_y^\star> \end{bmatrix} = \begin{bmatrix} I+Q & U+iV \\ U-iV & I-Q \end{bmatrix} = \Bmatrix
\end{equation}

@@

\section{Expanding our Jones Chain}
\label{section.RIME.JonesChain}
\pg
Having written our RIME as a matrix multiplication problem, we can now begin to differentiate between different types of Jones matrices. $\Jones_p$ describes the \emph{aggregate} effects which occur over the course of the signal's propagation to our antenna $p$. It can be split into a Jones chain of $n$ separate matrices, corresponding to different effects:
\begin{align}
\displaystyle \Jones_{sp}^H &= \prod_1^n \Jones_{nsp}^H = \Jones_{1sp}^H \Jones_{2sp}^H ... \Jones_{nsp}^H\\
\displaystyle \Jones_{sp}   &= \Jones_{nsp}...\Jones_{2sp} \Jones_{1sp}
\end{align}

\pg
Since the Jones terms are applied sequentially to the brightness matrix, starting with $\Jones_{1sp}$, the terms to the right of our decomposition of $\Jones_{sp}$ can be said to occur `at the source', i.e. before those to the left (said to occur `at the antenna').

@@

\pg
The physical quantity measured by an interferometric array, the \emph{visibility}, is not a function of the absolute phase of our signal, but rather a function of the \emph{phase difference} between our measured voltages $\volt_p$ and $\volt_q$. In other words, it is a function of the phase difference on \emph{baseline} $pq$, which consists of antennas $p$ and $q$. The direction which minimises this difference is called the \emph{phase centre}.

\pg
We shall use the conventional coordinate system and notations \citepads[][etc]{2011A&A...527A.106S,2001isra.book.....T}. We thus have a $z$ axis pointing towards the phase centre, and an antenna $p$ at coordinates $\U_p=(u_p,v_p,w_p)$. The phase difference between $p$ and phase centre (where, by definition, $\U=0$) for a signal arriving from direction $\pmb{\sigma}$ is then:
\begin{equation}
k_p=2\pi i (( u_p l + v_p m + w_p (n-1) ))
\end{equation}

@@

\pg
So far, $\Bmatrix$ represents the signal from a single point source. However, the formalism can be extended to multiple point sources. This is because electric fields -- and therefore radio signals -- are additive. If our sky contains $s$ point sources, then our RIME becomes:
\begin{equation}
\Vis_{pq} = \sum_s \Jones_{sp} \Bmatrix_s \Jones_{sq}^\Herm
\end{equation}

\pg
We now expand our Jones chain as before. Some, but not all, elements in the chain will only depend on the antenna. These will be the same for all sources. We can thus write our Jones chain as:
\begin{equation}
\Jones_{sp} = G_p E_{sp} K_{sp}
\end{equation}

@@

\pg
This is a three-dimensional Fourier transform. It can be reduced to a two-dimensional Fourier transform by writing the $n$-dependence in the exponential (and normalisation) as a Jones matrix, thus removing the \emph{non-coplanarity} terms from the explicit RIME. We can thus define the following:
\begin{align}
\Wjones_p &= \frac{1}{\sqrt{n}}e^{-2\pi i w_p(n-1)}\\
B_{pq}    &= \Ejones_p \Wjones_p \Bmatrix \Wjones_q^H \Ejones_q^h
\end{align}
and write our RIME in its final form:
\begin{equation}\label{eq.vis.FT.coherency.matrix}
\Vis_{pq} = \Gjones_p \bigg(\int_{\lvec} \Bmatrix_{pq}(\lvec) e^{-2\pi i\big(u_{pq}l+v_{pq}m+w_{pq}(n-1)\big)} dl dm \bigg) \Gjones_q^H
\end{equation}
\pg
Note that $\Bmatrix$ was the actual sky  brightness distribution as a function of direction, but now $\Bmatrix_{pq}$ is the sky brightness \emph{as seen by baseline pq} - there is no \emph{a priori} reason this should be the same for all baselines.\\

@@

\pg
Earlier calibration models and algorithm rely on the hypothesis that all baselines see the same sky. This is the \emph{fundamental premise of self-calibration}. The topic of self-cal is covered in \cref{section.calibration.2gc}. This premise only holds when all Direction-Dependent Effects (henceforth referred to as DDEs) are identical for all baselines (and therefore for all antennas). This is a necessary condition for the apparent sky $\Bmatrix_{pq}$ to be the same for all baselines\footnote{$\Bmatrix_{pq}$ would then correspond to the `true' sky attenuated by the ``power beam", which corresponds to our instrument's directional sensitivity in the traditional view}, as it allows us to write:
\begin{align}
\Ejones_p(\lvec) \Wjones_p(\lvec)&= \Ejones(\lvec) \Wjones(\lvec)\\
\Bmatrix_{pq}(\lvec) = \Bmatrix_{app}(\lvec) &= \Ejones(\lvec) \Wjones(\lvec) \Bmatrix)(\lvec) \Wjones^H(\lvec) \Ejones^H(\lvec)
\end{align}

@@

\pg
Note that deriving the van Cittert-Zernike theorem (\citepads{1934Phy.....1..201V}, covered in \citepads{2001isra.book.....T}) from the RIME was simply of matter of treating phase as a Jones matrix in its own right\footnote{\href{https://raw.githubusercontent.com/wiki/ska-sa/meqtrees/aips++_note185.pdf}{Noordm, J.E. 1996, AIPS++, Note 185: The Measurement Equation of a Generic Radio Telescope}}. A significant consequence of this is that DDEs can be incorporated within the RIME formalism.

\section{Time-dependent Jones matrices, Smearing, and Decoherence}
\label{section.RIME.TimeDep}

@@

\pg
The above means that direction-dependent effects must be time-independent for \cref{eq.fullsky.RIME} to hold. We can thus split direction-dependent effects into two categories: \emph{trivial} DDEs, which satisfy this condition (and can thus be solved for within the RIME formalism), and \emph{non-trivial} DDEs, which are - as the name indicates - non-trivial to account for within the RIME formalism. An example of a trivial DDE is the primary gain; and example of a non-trivial (i.e. time-variable) DDE is ionospheric turbulence.

\subsection{Smearing and Decoherence}
\label{section.RIME.TimeDep.Decoherence}

@@

\pg
The general effect is referred to as \emph{decoherence}, and the specific case of decoherence caused by the $\Kjones$ term is referred to as \emph{smearing}.

@@

\pg
The RIME presented thus far has been for an ideal case, free of noise in our measured visibility. In practice, each complex visibility will be affected by uncorrelated Gaussian noise\footnote{Detailed treatment given in \citepads{2001isra.book.....T}, section 6.2}, which means that a RIME of the same form as \cref{eq.multiple.source.RIME} would have to be formuled at follows:
\begin{equation}
\Vis_{pq} = \Gjones_p \bigg( \sum_s \Ejones_{sp} \Kjones_{sp} \Bmatrix_{s} \Kjones_{sq}^H \Ejones_{sq}^H \bigg) \Gjones_q + \sigma_{pq}
\end{equation}
where $\sigma_{pq}$ is a $(2,2)$ matrix with real and complex parts in each component. A similar addition can be made for any of the RIMEs presented here.

>>>>

\section{The RIME Formalism}\label{section.RIME}
%\minitoc

@@

\pg
Consider a single quasi-monochromatic point source in the sky. Its signal at a point in time and space can be described by a complex 2-vector $\evector$ -- this assumes plane waves. We can then represent physical processes which affect this signal's propagation using \emph{Jones matrices}. In other words:
\begin{equation}
\volt = \Jones\evector
\end{equation}
where $\volt$ is the voltage measured by our antenna, and $\Jones$ the Jones matrix describing the \emph{net} propagation effects - atmospheric, astrophysical, and instrumental - which affect our source's signal. $\Jones$ can be written as a series of matrix products, each individual matrix describing a physical phenomenon. They are then called a \emph{Jones chain}, and the resulting $\Jones$ is often called the \emph{total Jones matrix}.

\pg
A visibility is the correlation of voltage from two antennas. In other words,
\begin{align}
\Vis_{pq} &= 2 \langle\volt_p (\volt_q)^\Herm\rangle\\
\Vis_{pq} &= 2 \langle\Jones_p \evector (\Jones_q \evector)^\Herm\rangle\\
\Vis_{pq} &= 2 \langle\Jones_p \evector \evector^\Herm \Jones_q^\Herm\rangle
\end{align}
where $\Jones_p$ corresponds to the signal propagation chain of antenna $p$, and $\Jones_q$ for antenna $q$. Note that a factor of 2 has been introduced here, as a matter of convention. This is due to the definition of the Stokes parameter in relation to our $\evector$ outer product.

\pg
For $\mathbf{e}$ to be a 2-vector, rather than a 3-vecotr, we must choose to use $xyz$ as our coordinates, with $z$ the direction of propagation of our signal. \citetads{1996A&AS..117..137H} then show that the following relation holds:
\begin{equation}
2 \evector \evector^\Herm = 2  \begin{bmatrix} <e_x e_x^\star> & <e_x e_y^\star> \\ <e_y e_x^\star> & <e_y e_y^\star> \end{bmatrix} = \begin{bmatrix} I+Q & U+iV \\ U-iV & I-Q \end{bmatrix} = \Bmatrix
\end{equation}

@@

\section{Expanding our Jones Chain}
\label{section.RIME.JonesChain}
\pg
Having written our RIME as a matrix multiplication problem, we can now begin to differentiate between different types of Jones matrices. $\Jones_p$ describes the \emph{aggregate} effects which occur over the course of the signal's propagation to our antenna $p$. It can be split into a Jones chain of $n$ separate matrices, corresponding to different effects:
\begin{align}
\displaystyle \Jones_{sp}^H &= \prod_1^n \Jones_{nsp}^H = \Jones_{1sp}^H \Jones_{2sp}^H ... \Jones_{nsp}^H\\
\displaystyle \Jones_{sp}   &= \Jones_{nsp}...\Jones_{2sp} \Jones_{1sp}
\end{align}
where $s$ is a particular direction in the sky: these Jones matrices are therefore direction-dependent.
Since the Jones terms are applied sequentially to the brightness matrix, starting with $\Jones_{1sp}$, the terms to the right of our decomposition of $\Jones_{sp}$ can be said to occur `at the source', i.e. before those to the left (said to occur `at the antenna').

@@

\pg
The physical quantity measured by an interferometric array, the \emph{visibility}, is not a function of the absolute phase of our signal, but rather a function of the \emph{phase difference} between our measured voltages $\volt_p$ and $\volt_q$. In other words, it is a function of the phase difference on \emph{baseline} $pq$, which consists of antennas $p$ and $q$. The direction which minimises this difference is called the \emph{phase centre}\footnote{Bear in mind the limitations of this, described in \cref{sec.visibility}.}.

\pg
We shall use the conventional coordinate system and notations \citepads[][etc]{2011A&A...527A.106S,2001isra.book.....T}. We thus have a $z$ axis pointing towards the phase centre, and an antenna $p$ at coordinates $\U_p=(u_p,v_p,w_p)$. The phase difference between $p$ and phase centre (where, by definition, $\U=0$) for a signal arriving from direction $\pmb{\sigma}$ is then, assuming a small field of view:
\begin{equation}
k_p=2\pi i (( u_p l + v_p m + w_p (n-1) ))
\end{equation}

@@

\pg
So far, $\Bmatrix$ represents the signal from a single point source. However, the formalism can be extended to multiple point sources. This is because electric fields -- and therefore radio signals -- are additive, and that signal from different sources is not coherent (see \cref{eq.spatial.incor}), which means that visibilities are additive. If our sky contains $s$ point sources, then our RIME becomes:
\begin{equation}
\Vis_{pq} = \sum_s \Jones_{sp} \Bmatrix_s \Jones_{sq}^\Herm
\end{equation}

\pg
We now expand our Jones chain as before. Some, but not all, elements in the chain will only depend on the antenna. These will be the same for all sources. We can thus write our Jones chain as:
\begin{equation}
\Jones_{sp} = \Gjones_p \Ejones_{sp} \Kjones_{sp}
\end{equation}

@@

\pg
This is a three-dimensional Fourier transform. It can be reduced to a two-dimensional Fourier transform by writing the $n$-dependence in the exponential (and normalisation) as a Jones matrix, thus removing the \emph{non-coplanarity} terms from the explicit RIME. We can thus define the following:
\begin{align}
\Wjones_p &= \frac{1}{\sqrt{n}}e^{-2\pi i w_p(n-1)}\\
B_{pq}    &= \Ejones_p \Wjones_p \Bmatrix \Wjones_q^H \Ejones_q^h
\end{align}
and write our RIME in its final form:
\begin{equation}\label{eq.vis.FT.coherency.matrix}
\Vis_{pq} = \Gjones_p \bigg(\int_{\lvec} \Bmatrix_{pq}(\lvec) e^{-2\pi i\big(u_{pq}l+v_{pq}m+w_{pq}(n-1)\big)} dl dm \bigg) \Gjones_q^H
\end{equation}
\pg
Note that $\Bmatrix$ was the actual sky  brightness distribution as a function of direction, but now $\Bmatrix_{pq}$ is the sky brightness \emph{as seen by baseline pq} -- in the presence of direction-dependent effects, this is not the same for all baselines.\\

@@

\pg
Earlier calibration models and algorithms rely on the hypothesis that all baselines see the same sky. This is the \emph{fundamental premise of self-calibration}. The topic of self-cal is covered in \cref{section.calibration.2gc}. This premise only holds when all Direction-Dependent Effects (henceforth referred to as DDEs) are identical for all baselines (and therefore for all antennas). This is a necessary condition for the apparent sky $\Bmatrix_{pq}$ to be the same for all baselines\footnote{$\Bmatrix_{pq}$ would then correspond to the `true' sky attenuated by the ``power beam", which corresponds to our instrument's directional sensitivity in the traditional view}, as it allows us to write:
\begin{align}
\Ejones_p(\lvec) \Wjones_p(\lvec)&= \Ejones(\lvec) \Wjones(\lvec)\\
\Bmatrix_{pq}(\lvec) = \Bmatrix_{app}(\lvec) &= \Ejones(\lvec) \Wjones(\lvec) \Bmatrix)(\lvec) \Wjones^H(\lvec) \Ejones^H(\lvec)
\end{align}

@@

\pg
Note that deriving the van Cittert-Zernike theorem (\citepads{1934Phy.....1..201V}, covered in \citepads{2001isra.book.....T}) from the RIME was simply of matter of treating phase as a Jones matrix in its own right\footnote{\href{https://raw.githubusercontent.com/wiki/ska-sa/meqtrees/aips++_note185.pdf}{Noordam, J.E. 1996, AIPS++, Note 185: The Measurement Equation of a Generic Radio Telescope}}. A significant consequence of this is that DDEs can be incorporated within the RIME formalism.

\section{Time-dependent Jones matrices, Smearing, and Decorrelation}
\label{section.RIME.TimeDep}

@@

\pg
The above means that direction-dependent effects must be time-independent for \cref{eq.fullsky.RIME} to hold. We can thus split direction-dependent effects into two categories: \emph{trivial} DDEs, which satisfy this condition and are a simple multiplicative correction, and \emph{non-trivial} DDEs, which are not. An example of a trivial DDE is the primary gain; and example of a non-trivial (i.e. time-variable) DDE is ionospheric turbulence.

\subsection{Smearing and Decorrelation}
\label{section.RIME.TimeDep.Decoherence}

@@

\pg
The general effect is referred to as \emph{decorrelation}\footnote{Decorrelation can equivalently be referred to as ``decoherence".}, and the specific case of decoherence caused by the $\Kjones$ term is referred to as \emph{smearing}.

@@

\pg
The RIME presented thus far has been for an ideal case, free of noise in our measured visibility. In practice, each complex visibility will be affected by uncorrelated Gaussian noise\footnote{Detailed treatment given in \citepads{2001isra.book.....T}, section 6.2, which shows how to obtain the ideal radiometer equation}, which means that a RIME of the same form as \cref{eq.multiple.source.RIME} would have to be formuled at follows:
\begin{equation}
\Vis_{pq} = \Gjones_p \bigg( \sum_s \Ejones_{sp} \Kjones_{sp} \Bmatrix_{s} \Kjones_{sq}^H \Ejones_{sq}^H \bigg) \Gjones_q^H + \sigma_{pq}
\end{equation}
where $\sigma_{pq}$ is a $(2,2)$ matrix with real and complex parts in each component. A similar addition can be made for any of the RIMEs presented here.

### 2018-07-12 source/Ch2-techniques/calibration.tex
added various bits that came about after reshuffling the organisation

\section{Calibration Methods in Radio Interferometry}\label{section.calibration}
\pg
In this section, we will discuss the implementation of interferometric array calibration. Our analysis is based on the RIME formalism, described in \cref{section.RIME}. One key metric of calibration quality is the \emph{dynamic range}, mentioned in Sec. \ref{section.clean}. High dynamic ranges mean that a high contrast has been obtained, and fainter sources can be reached. Here, we redefine dynamic range as follows
\begin{equation}\label{eq.DR}
DR = \frac{flux_{max}}{\max(\sigma_{thermal},\sigma_{artefacts})}
\end{equation}
where $flux_{max}$ is the flux of the brightest source, $\sigma_{thermal}$ the thermal noise in the image, and $\sigma_{artefacts}$ the noise associated with calibration artefacts. This gives a definition for dynamic range useful even in fields with a single source. We can thus use it as a metric for calibration quality specifically.

@@

\pg
We can distinguish three `generations' of calibration methods, of increasing complexity. We will describe them in terms of the RIME, showing how each generation increases in generality to account for more exotic effects. They are referred to interchangeably as `nth-generation calibration' or `nGC' methods.

@@

\pg
First-generation calibration methods (1GC methods) consist of open-loop calibration. This relies entirely on instrument stability, and thus imposes significant design constraints on radio telescopes. It consists of briefly observing an external calibrator before and after each observation run to find a gain factor and offset error\footnote{For a concrete example, see \href{http://www.analog.com/en/analog-dialogue/articles/open-loop-calibration-techniques.html}{the open-loop calibration techniques page of analog.com}.}

\pg
Phase calibration in the 1GC era `proper' was not necessary, as engineers were capable of ensuring adequate phase stability in contemporary interferometers. Phase was thus calculated relative to a fixed frame of reference, usually the central antenna of a 3-antenna array.

\pg
In RIME terms, this consists of solving for a very basic form of $\Gjones_p$:
\begin{equation}
\Gjones_p = \begin{bmatrix} a_p t + b_p \end{bmatrix} % \begin{bmatrix} e^{2 \pi i \nu ( \phi_p - \phi_0)} \end{bmatrix}
\end{equation}
where $a_p,b_p$ are constants solved for during open-loop calibration and $t$ is time. %, $\nu$ the observing frequency, $\phi$ the phase at an antenna, and $\phi_0$ the phase at the reference antenna.

@@

\pg
Self-calibration and adaptive optics are the same thing, albeit meeting different constraints. Most notably, interferometric data is digital, allowing radio astronomers to perform their adaptive optics correction after the observation.

\pg
If amplitude and phase gains can be written as antenna-dependent, then each antenna-based error is estimated N times (once per baseline which includes this antenna). By estimating, and correcting for, these errors, one can use a simple source model to infer an improved one. This is why the method is referred to as self-cal; by calibrating `on' a good calibrator source (bright, compact and unresolved), one can drastically improve one's source model, along with one's calibration solutions.

>>>>

\section{Calibration Methods in Radio Interferometry}\label{section.calibration}
\pg
In this section, we will discuss the implementation of interferometric array calibration. Our analysis is based on the RIME formalism, described in \cref{section.RIME}. One key metric of calibration quality is the \emph{dynamic range}, mentioned in Sec. \ref{section.clean}. High dynamic ranges mean that a high contrast has been obtained, and fainter sources can be reached. Here, because the difference between thermal noise and artefacts becomes relevant, we redefine dynamic range as follows
\begin{equation}\label{eq.DR}
DR = \frac{flux_{max}}{S_{min},\max(\sigma_{thermal},\sigma_{artefacts})}
\end{equation}
where $S_{max}$ is the flux of the brightest source, $S_{min}$ the flux of the faintest source detectable above the noise, $\sigma_{thermal}$ the thermal noise in the image, and $\sigma_{artefacts}$ the noise associated with calibration artefacts. This gives a definition for dynamic range useful even in fields with a single source. We can thus use it as a metric for calibration quality specifically.

@@

\pg
We can distinguish three `generations' of calibration methods, of increasing complexity \citepads{2010A&A...524A..61N}. We will describe them in terms of the RIME, showing how each generation increases in generality to account for more exotic effects. They are referred to interchangeably as `nth-generation calibration' or `nGC' methods.

@@

\pg
First-generation calibration methods (1GC methods) consist of open-loop calibration. This relies entirely on instrument stability, and thus imposes significant design constraints on radio telescopes. It consists of briefly observing an external calibrator before and after each observation run to find calibration solutions for those calibrator observations, and then interpolate between them over the ``target" observation. %\footnote{For a concrete example, see \href{http://www.analog.com/en/analog-dialogue/articles/open-loop-calibration-techniques.html}{the open-loop calibration techniques page of analog.com}.}
%\pg
%Phase calibration in the 1GC era `proper' was not necessary, as engineers were capable of ensuring adequate phase stability in contemporary interferometers. Phase was thus calculated relative to a fixed frame of reference, usually the central antenna of a 3-antenna array. 
In RIME terms (see \cref{section.RIME}), this consists of solving for a very basic form of $\Gjones_p$:
\begin{equation}
\Gjones_p = \begin{bmatrix} a_p t + b_p \end{bmatrix} % \begin{bmatrix} e^{2 \pi i \nu ( \phi_p - \phi_0)} \end{bmatrix}
\end{equation}
where $a_p,b_p$ are constants solved for during open-loop calibration and $t$ is time. %, $\nu$ the observing frequency, $\phi$ the phase at an antenna, and $\phi_0$ the phase at the reference antenna.

@@

\pg
Self-calibration and adaptive optics are essentially the same thing, albeit meeting different constraints. Most notably, interferometric data is digital, allowing radio astronomers to perform their adaptive optics correction after the observation.

\pg
If amplitude and phase gains can be written as antenna-dependent, then each antenna-based error is estimated N times (once per each baseline which includes this antenna). By estimating, and correcting for, these errors, one can use a simple source model to infer an improved one. This is why the method is referred to as self-cal; by calibrating `on' a good calibrator source (bright, compact and unresolved), one can drastically improve one's source model, along with one's calibration solutions.
